Search, Compress, Compile: Library Learning in Neurally-Guided Bayesian Program Learning

### 2018-05-18 nips_2018.tex
Follows (some) chktex recommandations

Because any computable learning problem can in principle be cast as
program induction, it is important to delimit our focus.  In contrast
to computer assisted programming~\cite{solar2008program} or genetic
programming~\cite{DBLP:books/daglib/0070933}, our goal is not to automate software
engineering, to learn to synthesize large bodies of code, or to learn
complex programs starting from scratch.  Ours is a basic AI goal:
capturing the human ability to learn to think flexibly and efficiently
in new domains -- to learn what you need to know about a domain so you
don't have to solve new problems starting from scratch.  We are
focused on the kinds of problems that humans can solve relatively
quickly, once they acquire the relevant domain expertise.  These
correspond to tasks solved by short programs -- if you have an
expressive DSL.  Even with a good DSL, program search may be
intractable;
%
% , and adding new library routines to the DSL only broadens the
% search space;
%
so we amortize the cost of search by training a
neural network to assist the search procedure.

@@

The learned DSL effectively encodes a prior on programs likely to
solve tasks in the domain, while the neural net looks at the example
input-output pairs for a specific task and produces a ``posterior''
for programs likely to solve that specific task.  The neural network
thus functions as a \textbf{recognition model} supporting a form of
approximate Bayesian program induction, jointly trained with a
\textbf{generative model} for programs encoded in the DSL, in the
spirit of the Helmholtz machine ~\cite{hinton1995wake}). The
recognition model ensures that searching for programs remains
tractable even as the DSL (and hence the search space for programs)
expands.

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference  by maximizing this lower bound on the joint probability,
alternating maximization w.r.t. the frontiers (Search) and the DSL (Compression):
\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t. the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

Searching for programs is hard because
of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $q(\cdot |\cdot )$,
during the compilation phase: $q$ is trained to approximate the
posterior over programs, $q(p|x)\approx \probability[p|x,\mathcal{D},\theta}]\propto\probability[x|p]\probability[p|\mathcal{D},\theta}]$,
  thus amortizing the cost of finding programs with high posterior probability.

@@

\end{tikzpicture}
  \end{minipage}
\hspace{0.1cm}\begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (- (length $\ell$) 1))))}\\
      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (+ 1 1))))}\\
    \end{tabular}&
    \begin{tabular}{l}
      \code{(map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{(map)}}\code{(range $\alpha$))}\\
    \end{tabular}\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
      \end{tabular}&
    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x $\alpha$) $\beta$ x))) s)}
      \end{tabular}
\\\bottomrule\\
\end{tabular}
\caption{\textbf{Left:} syntax trees of two programs sharing common structure, highlighted in {\color{red} red},
  from which we extract a fragment and add it to the DSL (bottom). \textbf{Right:} actual programs,
  from which we extract fragments that  (top) slice from the beginning of a list or (bottom) perform character substitutions.}

  \label{fragmentExample}
\end{figure}

@@

\begin{wrapfigure}{r}{0.55\textwidth}\centering
  \begin{tabular}{lll}
    \toprule
    Name & Input & Output \\\midrule
    repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
    keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
    product & [7\, 1\, 6\, 2] & 84 \\
    \bottomrule
  \end{tabular}
  \captionof{table}{Some tasks in our list function domain. See the
  supplement for the complete data set.}
  \label{listexamples}
\end{wrapfigure}
We created 236 human-interpretable list manipulation tasks, each with 15
input/output examples (Tbl.~\ref{listexamples}).
Our data set is interesting in three major ways: many of the tasks require
complex solutions; the tasks were not generated from some latent DSL, and
the agent must learn to solve these complicated problems from only 236
tasks.
Our data set assumes arithmetic operations as well as sequence operations,
so we additionally provide our system with the following arithmetic
primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
\code{is-prime}.

@@

Because our enumerative search procedure cannot generate string
constants, we instead enumerate programs with string-valued
parameters.  For example, to learn a program that prepends ``Dr.'', we
enumerate $\text{\code{(}}f_3\code{ string s)}$ -- where $f_3$ is the
learned appending primitive (Fig.~\ref{initialExampleDSL}) -- and then
define $\probability[x|p]$ by approximately marginalizing out the
string parameters via a simple dynamic program.
In Sec.~\ref{regressionSection}, we will use a similar trick to
synthesize programs containing real numbers, but using gradient
descent instead of dynamic programming.

@@

\section{Symbolic Regression: Programs from visual input}\label{regressionSection}
We apply \system
to symbolic regression problems.  Here, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
100 symbolic regression problems, each either a polynomial of degree
1--4 or a rational function.  The recognition model is a
convolutional network that observes an image of the target function's
graph (Fig.~\ref{functions}) -- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the recognition model can learn to look at a graph and predict
what kind of function best explains it.  A key difficulty, however, is
that these problems are best solved with programs containing real
numbers.  Our solution to this difficulty is to enumerate
 programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.
We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.

\begin{wrapfigure}{r}{0.45\textwidth}\vspace{-0.5cm} \newcommand{\functionSize}{1.5cm}\centering
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/48.png}
  \includegraphics[width = \functionSize]{figures/functions/102.png}
  \includegraphics[width = \functionSize]{figures/functions/116.png}\\
  \vspace{1pt}\includegraphics[width = \functionSize]{figures/functions/181.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/149.png}
    \includegraphics[width = \functionSize]{figures/functions/187.png}
  \caption{Recognition model input for symbolic regression. DSL learns subroutines for polynomials (top row) and rational functions (bottom row) while the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines best explains the observation.}\label{functions}\vspace{-2cm}
\end{wrapfigure}
\system learns a DSL containing 13 new functions,
most of which are templates for polynomials of different orders or ratios of polynomials.
It also learns to find programs that minimize the number of continuous degrees of freedom.
For example, it learns to represent linear functions with the program
\code{(* real (+ x real))}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing -- both the implicit bias towards shorter programs and the likelihood model's BIC penalty.

@@

We compare with ablations of our model on held out tasks.
The purpose of this ablation study is 
both to examine the role of each component of \systemEnding,
as well as to compare with
prior approaches in the literature:
a head-to-head
comparison of program synthesizers is complicated by the fact that
each system, including ours, makes idiosyncratic 
assumptions about the space of programs and the statement of tasks.
\begin{wraptable}{r}{0.55\textwidth}
\vspace{-5pt}
%% This radically shrinks the table
\tabcolsep=2.5pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{lcccccc}
  \toprule& Ours& 
    No NN & SE&NPS & PCFG & Enum \\
  \midrule
  \multicolumn{7}{c}{\emph{List Processing}}\\
  \midrule
  \% solved&\textbf{79\%}\footnotemark & 
  76\% &71\%&35\%&62\%&37\%\\
  Solve time&  4.1s&5.8s&10.6s&34.7s&43.4s&20.2s\\
  \midrule
  \multicolumn{7}{c}{\emph{Text Editing}}\\
  \midrule
  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
  Solve time& 29s&65s&38s&80s&--&235s\\
  \midrule
  \multicolumn{7}{c}{\emph{Symbolic Regression}}\\
  \midrule
  \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
  Solve time&  24s& 40s  &28s&31s&55s&29s\\
  \bottomrule
  \end{tabular}
\caption{\% solved before timeout. Solve time: averaged over solved
  tasks.  %SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
  %structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
Nevertheless, much prior work can be modeled within our setup. 
We compare with the following ablations (Table~\ref{baselineComparisons}):
\\\noindent \textbf{No NN:} lesions the recognition model.
\\\noindent \textbf{NPS}, which does not learn the DSL,
instead learning the recognition model
from samples drawn from the fixed DSL.
We call this NPS (Neural Program Synthesis)
because this is closest to how
RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL. This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
\\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
but instead learns the parameters of the DSL ($\theta$) -- learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first search step.
\footnotetext[2]{Due to a last-minute bug we ran the list processing experiments without training the recognition model on ``Helmholtz machine'' style data. We are rerunning this experiment and anticipate improved quantitative results for list processing -- notice the recognition model helps much more for text editing and symbolic regression.} &
%For each domain,
%% We are interested both in how many tasks the
%% agent can solve and how quickly it can find those solutions.
%% Tbl.~\ref{baselineComparisons}
%% compares our model against these alternatives.
%% We consistently
%% improve on the baselines,
%% and find that lesioning the recognition model
%% and lesioning it also slows down the convergence of the algorithm,
%% taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
%% This supports a view of the recognition model as a way of amortizing the cost of search.

\begin{figure}\centering
  \includegraphics[width = 4.5cm]{figures/list.eps}
  \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}        
  \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}   
  \caption{Learning curves for \system both with (\orange{orange}) and without
  (\teal{teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves} 
\end{figure}

@@

%\paragraph{Outlook.}
We contribute an algorithm, \systemEnding, that learns to program by
bootstrapping a DSL with new domain-specific primitives that the algorithm
itself discovers, together with a neural recognition model that learns how to
efficiently deploy the DSL on new tasks. We believe this integration of top-down
symbolic representations and bottom-up neural networks --- both of them learned
--- helps make program induction systems more generally useful for AI. Many
directions remain open.
%\paragraph{Future.}
Two immediate goals are to integrate more sophisticated neural recognition
models~\cite{devlin2017robustfill} and program
synthesizers~\cite{solar2008program}, which may improve performance in some
domains over the generic methods used here.
We are in the process of applying our algorithm to generative programs and
prototyped this with a turtle-like domain: Figure \ref{geomCompiled} gives some
preliminary results for turtle graphics  --- see Section 1 of the supplementary material for more
details.  Another direction is to explore DSL meta-learning: can we find a
\emph{single} universal primitive set that could effectively bootstrap DSLs for
new domains, including the three domains considered,  but also many others?
}\hfill
\begin{minipage}{0.4\textwidth}
  \centering
  \includegraphics[width=\textwidth]{figures/geomCompiled.eps} 
  \captionof{figure}{
  Top left: Example training tasks. Top right: samples from the learned DSL.
 Bottom: \% holdout testing tasks solved (middle), on sides are the averaged samples from the DSL
  before any training (left) and after last iteration (right).}
\label{geomCompiled}
\end{minipage}

>>>>

Because any computable learning problem can in principle be cast as
program induction, it is important to delimit our focus.  In contrast
to computer assisted programming~\cite{solar2008program} or genetic
programming~\cite{DBLP:books/daglib/0070933}, our goal is not to automate software
engineering, to learn to synthesize large bodies of code, or to learn
complex programs starting from scratch.  Ours is a basic AI goal:
capturing the human ability to learn to think flexibly and efficiently
in new domains --- to learn what you need to know about a domain so you
don't have to solve new problems starting from scratch.  We are
focused on the kinds of problems that humans can solve relatively
quickly, once they acquire the relevant domain expertise.  These
correspond to tasks solved by short programs --- if you have an
expressive DSL.  Even with a good DSL, program search may be
intractable;
%
% , and adding new library routines to the DSL only broadens the
% search space;
%
so we amortize the cost of search by training a
neural network to assist the search procedure.

@@

The learned DSL effectively encodes a prior on programs likely to
solve tasks in the domain, while the neural net looks at the example
input-output pairs for a specific task and produces a ``posterior''
for programs likely to solve that specific task.  The neural network
thus functions as a \textbf{recognition model} supporting a form of
approximate Bayesian program induction, jointly trained with a
\textbf{generative model} for programs encoded in the DSL, in the
spirit of the Helmholtz machine~\cite{hinton1995wake}). The
recognition model ensures that searching for programs remains
tractable even as the DSL (and hence the search space for programs)
expands.

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference  by maximizing this lower bound on the joint probability,
alternating maximization w.r.t.\ the frontiers (Search) and the DSL (Compression):
\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

Searching for programs is hard because
of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $q(\cdot |\cdot )$,
during the compilation phase: $q$ is trained to approximate the
posterior over programs, $q(p|x)\approx \probability[p|x,\mathcal{D},\theta]\propto\probability[x|p]\probability[p|\mathcal{D},\theta}]$,
  thus amortizing the cost of finding programs with high posterior probability.

@@

\end{tikzpicture}
  \end{minipage}
\hspace{0.1cm}\begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (- (length $\ell$) 1))))}\\
      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (+ 1 1))))}\\
    \end{tabular}&
    \begin{tabular}{l}
      \code{(map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{(map)}}\code{(range $\alpha$))}\\
    \end{tabular}\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
      \end{tabular}&
    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x $\alpha$) $\beta$ x))) s)}
      \end{tabular}
\\\bottomrule\\
\end{tabular}
\caption{\textbf{Left:} syntax trees of two programs sharing common structure, highlighted in {\color{red} red},
  from which we extract a fragment and add it to the DSL (bottom). \textbf{Right:} actual programs,
  from which we extract fragments that  (top) slice from the beginning of a list or (bottom) perform character substitutions.}\label{fragmentExample}
\end{figure}

@@

\begin{wrapfigure}{r}{0.55\textwidth}\centering
  \begin{tabular}{lll}
    \toprule
    Name & Input & Output \\\midrule
    repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
    keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
    product & [7\, 1\, 6\, 2] & 84 \\
    \bottomrule
  \end{tabular}
  \captionof{table}{Some tasks in our list function domain. See the
  supplement for the complete data set.}\label{listexamples}
\end{wrapfigure}
We created 236 human-interpretable list manipulation tasks, each with 15
input/output examples (Tbl.~\ref{listexamples}).
Our data set is interesting in three major ways: many of the tasks require
complex solutions; the tasks were not generated from some latent DSL, and
the agent must learn to solve these complicated problems from only 236
tasks.
Our data set assumes arithmetic operations as well as sequence operations,
so we additionally provide our system with the following arithmetic
primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
\code{is-prime}.

@@

Because our enumerative search procedure cannot generate string
constants, we instead enumerate programs with string-valued
parameters.  For example, to learn a program that prepends ``Dr.'', we
enumerate $\text{\code{(}}f_3\code{ string s)}$ -- where $f_3$ is the
learned appending primitive (Fig.~\ref{initialExampleDSL}) --- and then
define $\probability[x|p]$ by approximately marginalizing out the
string parameters via a simple dynamic program.
In Sec.~\ref{regressionSection}, we will use a similar trick to
synthesize programs containing real numbers, but using gradient
descent instead of dynamic programming.

@@

\section{Symbolic Regression: Programs from visual input}\label{regressionSection}
We apply \system
to symbolic regression problems.  Here, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
100 symbolic regression problems, each either a polynomial of degree
1--4 or a rational function.  The recognition model is a
convolutional network that observes an image of the target function's
graph (Fig.~\ref{functions}) --- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the recognition model can learn to look at a graph and predict
what kind of function best explains it.  A key difficulty, however, is
that these problems are best solved with programs containing real
numbers.  Our solution to this difficulty is to enumerate
 programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.
We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.

\begin{wrapfigure}{r}{0.45\textwidth}\vspace{-0.5cm} \newcommand{\functionSize}{1.5cm}\centering
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/48.png}
  \includegraphics[width = \functionSize]{figures/functions/102.png}
  \includegraphics[width = \functionSize]{figures/functions/116.png}\\
  \vspace{1pt}\includegraphics[width = \functionSize]{figures/functions/181.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/149.png}
    \includegraphics[width = \functionSize]{figures/functions/187.png}
  \caption{Recognition model input for symbolic regression. DSL learns subroutines for polynomials (top row) and rational functions (bottom row) while the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines best explains the observation.}\label{functions}\vspace{-2cm}
\end{wrapfigure}
\system learns a DSL containing 13 new functions,
most of which are templates for polynomials of different orders or ratios of polynomials.
It also learns to find programs that minimize the number of continuous degrees of freedom.
For example, it learns to represent linear functions with the program
\code{(* real (+ x real))}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing --- both the implicit bias towards shorter programs and the likelihood model's BIC penalty.

@@

We compare with ablations of our model on held out tasks.
The purpose of this ablation study is 
both to examine the role of each component of \systemEnding,
as well as to compare with
prior approaches in the literature:
a head-to-head
comparison of program synthesizers is complicated by the fact that
each system, including ours, makes idiosyncratic 
assumptions about the space of programs and the statement of tasks.
\begin{wraptable}{r}{0.55\textwidth}
\vspace{-5pt}
%% This radically shrinks the table
\tabcolsep=2.5pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{lcccccc}
  \toprule& Ours& 
    No NN & SE&NPS & PCFG & Enum \\
  \midrule
  \multicolumn{7}{c}{\emph{List Processing}}\\
  \midrule
  \% solved&\textbf{79\%}\footnotemark & 
  76\% &71\%&35\%&62\%&37\%\\
  Solve time&  4.1s&5.8s&10.6s&34.7s&43.4s&20.2s\\
  \midrule
  \multicolumn{7}{c}{\emph{Text Editing}}\\
  \midrule
  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
  Solve time& 29s&65s&38s&80s&--&235s\\
  \midrule
  \multicolumn{7}{c}{\emph{Symbolic Regression}}\\
  \midrule
  \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
  Solve time&  24s& 40s  &28s&31s&55s&29s\\
  \bottomrule
  \end{tabular}
\caption{\% solved before timeout. Solve time: averaged over solved
  tasks.  %SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
  %structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
Nevertheless, much prior work can be modeled within our setup. 
We compare with the following ablations (Table~\ref{baselineComparisons}):
\\\noindent \textbf{No NN:} lesions the recognition model.
\\\noindent \textbf{NPS}, which does not learn the DSL,
instead learning the recognition model
from samples drawn from the fixed DSL.
We call this NPS (Neural Program Synthesis)
because this is closest to how
RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL. This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
\\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
but instead learns the parameters of the DSL ($\theta$) --- learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first search step.
\footnotetext[2]{Due to a last-minute bug we ran the list processing experiments without training the recognition model on ``Helmholtz machine'' style data. We are rerunning this experiment and anticipate improved quantitative results for list processing -- notice the recognition model helps much more for text editing and symbolic regression.} &
%For each domain,
%% We are interested both in how many tasks the
%% agent can solve and how quickly it can find those solutions.
%% Tbl.~\ref{baselineComparisons}
%% compares our model against these alternatives.
%% We consistently
%% improve on the baselines,
%% and find that lesioning the recognition model
%% and lesioning it also slows down the convergence of the algorithm,
%% taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
%% This supports a view of the recognition model as a way of amortizing the cost of search.

\begin{figure}\centering
  \includegraphics[width = 4.5cm]{figures/listLearningCurve.eps}
  \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}
  \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}
  \caption{Learning curves for \system both with (\orange{in orange}) and without
  (\teal{in teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves}
\end{figure}

@@

%\paragraph{Outlook.}
We contribute an algorithm, \systemEnding, that learns to program by
bootstrapping a DSL with new domain-specific primitives that the algorithm
itself discovers, together with a neural recognition model that learns how to
efficiently deploy the DSL on new tasks. We believe this integration of top-down
symbolic representations and bottom-up neural networks --- both of them learned
--- helps make program induction systems more generally useful for AI. Many
directions remain open.
%\paragraph{Future.}
Two immediate goals are to integrate more sophisticated neural recognition
models~\cite{devlin2017robustfill} and program
synthesizers~\cite{solar2008program}, which may improve performance in some
domains over the generic methods used here.
We are in the process of applying our algorithm to generative programs and
prototyped this with a turtle-like domain: Figure~\ref{geomCompiled} gives some
preliminary results for turtle graphics  --- see Section 1 of the supplementary material for more
details.  Another direction is to explore DSL meta-learning: can we find a
\emph{single} universal primitive set that could effectively bootstrap DSLs for
new domains, including the three domains considered,  but also many others?
}\hfill
\begin{minipage}{0.4\textwidth}
  \centering
  \includegraphics[width=\textwidth]{figures/geomCompiled.eps} 
  \captionof{figure}{
  Top left: Example training tasks. Top right: samples from the learned DSL.
 Bottom: \% holdout testing tasks solved (middle), on sides are the averaged samples from the DSL
  before any training (left) and after last iteration (right).}
\label{geomCompiled}
\end{minipage}
